### 2010-11-19 docs/tex/review.tex
Genetic Algorithms fairly completed.

\subsection{Evolutionary Algorithms}
Algorithms which are inspired by the strategy of \textit{survival of the fittest} from evolutionary biology, and uses different methods to optimize mathematical expressions from a possible pool of population of solutions, can be defined as evolutionary algorithms. There are many different forms of evolutionary algorithms. All of which can be generalized to follow a certain pattern. The mathematical expression, also termed as \textbf{fitness function} is used to evaluate the fitness of each individual of the set of existing solution space. An individual or a solution which provides higher output from the fitness function is considered as more fit to survive in the environment. So given a problem the algorithm initializes the environment with a random set of population or solutions. Then it uses the fitness function to select a better set of solutions from the existing once. After that different operators of \textbf{mutation} and \textbf{recombination} are applied to the selected set of solution to come up with a new generation of solution set or population born from the existing ones. Again the fitness function is applied to the current set of population to come up with a better set and this process iterates until the evaluation from the fitness function is satisfactory to give the best set of solutions.

@@

Using GA for a binary representation mutation can be single point or multiple point, where a random location is selected from the binary genome and the bit is flipped. For integer representation random resetting is used instead of bit flipping. Creep mutation is another scheme which was designed for ordinal attributes and works by adding a small (positive or negative) value to each gene with a random probability. For a real-valued or floating-point mutation none of the above strategy is applicable. As for this case mutation is applied to a random gene maintaining a certain range of value within an upper and a lower bound and using a standard for randomization, such as Uniform, Gaussian or Cauchy distribution. For permutation representation swap mutation is a strategy used by simply swapping the location of two randomly selected genes. Insert mutation is another strategy where a random gene is selected and inserted to a different location in the genome, causing the other genes to move a single space. In scramble mutation, location of a collection of randomly selected genes are scrambled, while in inversion mutation serialized location of a set of genes are inverted.

GA also have many different recombination operators for each representation. For binary and integer representation one-point crossover and \textit{N}-point-crossover are the most frequently used. These two operators work by dividing the parent genome at one or multiple random points and combining them to create an offspring. Uniform-crossover \cite{sywerda1989} is another method which works by treating each gene independently and making a random choice at to which parent it should inherit from. Recombination for floating-point representation can use the same operators like binary and integer, and when it does it is termed as discreet recombination. Another type is intermediate or arithmetic recombination where the offspring do not just receive a part of their parent genome but instead their genomes are derived from their using an arithmetic formula. Permutation representation has many interesting recombination operators out of which partially mapped crossover was first proposed by Goldsberg and Lingle \cite{goldberg1985} for the traveling salesman problem and since then it has become one of the most widely used operators for adjacency type problems. Other useful operators for permutation representation are Edge crossover, Order crossover and Cycle crossover, where their name provides more or less an appropriate description of their operation.

>>>>

\subsection{Evolutionary Algorithms}
Algorithms which are inspired by the strategy of \textit{survival of the fittest} from evolutionary biology, and uses different methods to optimize mathematical expressions which are used to evaluate a pool of population of solutions over generation, can be defined as evolutionary algorithms. There are many different forms of evolutionary algorithm. All of which can be generalized to follow their pattern from natural selection. The mathematical expression, also termed as \textbf{fitness function} is used to evaluate the fitness of each individual from the set of existing solution space. An individual or a solution which provides higher output from the fitness function is considered as more fit to survive in the environment. So given a problem the algorithm initializes the environment with a random set of population or solutions. Then it uses the fitness function to select a better set of solutions from the existing once. After that different operators of \textbf{mutation} and \textbf{recombination} are applied to the selected set of solution to come up with a new generation of solution set born from the existing ones. Again the fitness function is applied to the current set of population to come up with a better set and this process iterates until the evaluation from the fitness function is satisfactory to give the best set of solutions.

@@

Using GA for a binary representation mutation can be single point or multiple point, where a random location is selected from the binary genome and the bit is flipped. For integer representation random resetting is used instead of bit flipping. Creep mutation is another scheme which was designed for ordinal attributes and works by adding a small (positive or negative) value to each gene with a random probability. For a real-valued or floating-point mutation none of the above strategy is applicable. As for this case mutation is applied to a random gene maintaining a certain range of value within an upper and a lower bound and using a standard for randomization, such as Uniform, Gaussian or Cauchy distribution. For permutation representation, swap mutation is a strategy used by simply swapping the location of two randomly selected genes. Insert mutation is another strategy where a random gene is selected and inserted to a random location in the genome, causing the other genes to move a single space. In scramble mutation, location of a collection of randomly selected genes are scrambled, while in inversion mutation serialized location of a set of genes are inverted.

GA also have many different recombination operators for each representation. For binary and integer representation one-point crossover and \textit{N}-point-crossover are the most frequently used. These two operators work by dividing the parent genome at one or multiple random points and combining them to create an offspring. Uniform-crossover \cite{sywerda1989} is another method which works by treating each gene independently and making a random choice at to which parent it should inherit from. Recombination for floating-point representation can use the same operators like binary and integer, and when it does it is termed as discreet recombination. Another type is intermediate or arithmetic recombination where the offspring do not just receive a part of their parent genome but instead their genomes are derived from their parent using an arithmetic formula. Permutation representation has many interesting recombination operators, out of which partially mapped crossover was first proposed by Goldsberg and Lingle \cite{goldberg1985} for the traveling salesman problem and since then it has become one of the most widely used operators for adjacency type problems. Other useful operators for permutation representation are Edge crossover, Order crossover and Cycle crossover, where their name provides more or less an appropriate description of their operation.

In regards to parent selection the most commonly used process by GA is fitness proportion selection, where the selection probability of a parent depends on the absolute fitness value of the individual compared to the absolute fitness values of the rest of the population. A fitness value is evaluated from the evaluation or the fitness function. Introduced by Holland \cite{holland1975} this procedure has a few problems. Premature convergence being one where individuals that are a lot better than the rest take over the entire population. Also when fitness values are all very close there is almost no selection pressure in evolution of the fittest. Inspired by the observed drawbacks of fitness proportion selection a method was proposed by Baker \cite{baker1987} termed as Ranking selection. This procedure preserves a constant selection pressure by sorting the population on the basis on fitness, and then allocating selection probabilities to individuals according to their rank. Both of the above mentioned selection procedures requires knowledge of the entire population. But when the population size is very large Tournament selection is an effective process. In this case a randomized pre-selection is done over the large population and then they are ranked according to their fitness value and selected as parent.

Genetic Algorithms are one of the most effective strategy used in the field of evolutionary computation to solve optimization problems. Then again there are limitations. When the solution set of a problem do not have a gnomic representation, genetic algorithms cannot play any role in improving it. So the set of problems to which GA can be applied is not universal.
